### 2022-09-16 src/MS1_report01.Rmd
add report for milestone 1

---
title: "Milestone 1: Play with the data"
author: "Anders"
date: "6/9/2022"
output: pdf_document
#output: html_notebook
editor_options: 
  chunk_output_type: console
---

# Load R packages
```{r include=FALSE}
# install.packages("glmnet", repos = "https://cran.us.r-project.org")
library(MASS)
library(glmnet)
library(mboost)
library(pcalg)
```

## Read data "6 genes" and create dataframes and matrixes
### NAN is removed
```{r}
df01 <- readxl::read_xlsx('/Users/anders/Documents/Master/data/SCR_SUR_6genes_noNAN.xlsx')
df02 <- df01[df01$timepoint %in% "SCR", ]
df03 <- df01[df01$timepoint %in% "SUR", ]
df04 <- df02
df04$ProliferationScore <- df03$ProliferationScore
df05 <- subset(df04, TrialArmNeo %in% "Letro+Ribo", select=-c(UniqueID, timepoint, TrialArmNeo))

ma02 <- as.matrix(subset(df02, select=-c(UniqueID, timepoint, TrialArmNeo)))
ma03 <- as.matrix(subset(df03, select=-c(UniqueID, timepoint, TrialArmNeo)))
ma04 <- as.matrix(subset(df04, select=-c(UniqueID, timepoint, TrialArmNeo)))
ma05 <- as.matrix(subset(df05))
```

## Formula for modles
```{r}
fm01 = ProliferationScore ~ CCND1 + CCNE1 + CDKN1A + ESR1 + MYC + RB1
fm02 = ProliferationScore ~ CCND1*RB1 + CCND1*CDKN1A + CCND1*ESR1 + CCNE1*CDKN1A + CCNE1*RB1 + MYC*ESR1 + MYC*CDKN1A
fm03 = ProliferationScore ~ CCND1:RB1 + CCND1:CDKN1A + CCND1:ESR1 + CCNE1:CDKN1A + CCNE1:RB1 + MYC:ESR1 + MYC:CDKN1A
```

# A: Variable selection focusing on one timpoint at a time (SCR or SUR)

## plot at SCR
```{r echo=FALSE}
par(mfrow=c(2,3), mar=rep(2,4))
names = c('CCND1', 'CCNE1', 'CDKN1A', 'ESR1', 'MYC','RB1') 
for(i in 1:6){
  x = ma02[,i+1]
  y = ma02[,1]
  plot(x, y, main=names[i])
  abline(lm(y ~ x), col = "red", lwd=3)
  mtext(paste("Correlation:", round(cor(x, y), 2)), side=1, line=-1.5, cex=1.2)
  cat(names[i])
  print(cor(x,y), method = c("pearson"))
}
```

## plot at SUR
```{r echo=FALSE}
par(mfrow=c(2,3), mar=rep(2,4))
names = c('CCND1', 'CCNE1', 'CDKN1A', 'ESR1', 'MYC','RB1') 
for(i in 1:6){
  x = ma03[,i+1]
  y = ma03[,1]
  plot(x, y, main=names[i])
  abline(lm(y ~ x), col = "red", lwd=3)
  mtext(paste("Correlation:", round(cor(x, y), 2)), side=1, line=-1.5, cex=1.2)
  cat(names[i])
  print(cor(x,y), method = c("pearson"))
}
```

## plot prolif: SCR vs SUR
```{r echo=FALSE}
par(mfrow=c(1,1))
x=df02$ProliferationScore; y=df03$ProliferationScore
plot(x, y)
abline(lm(y ~ x), col = "red", lwd=3)
mtext(paste("Correlation:", round(cor(x, y), 2)), side=1, line=-1.5, cex=1.2)
```

## plot SCR genes vs SUR Proliferation
### But: the two treatments are pooled 
```{r echo=FALSE}
par(mfrow=c(2,3), mar=rep(2,4))
names = c('CCND1', 'CCNE1', 'CDKN1A', 'ESR1', 'MYC','RB1') 
for(i in 1:6){
  x = ma02[,i+1]
  y = ma03[,1]
  plot(x, y, main=names[i])
  abline(lm(y ~ x), col = "red", lwd=3)
  mtext(paste("Correlation:", round(cor(x, y), 2)), side=1, line=-1.5, cex=1.2)
  cat(names[i])
  print(cor(x,y), method = c("pearson"))
}
```

## SCR: Step AIC/BIC
```{r}
lm01 = lm(fm01, data = df02)
lm02 = lm(fm02, data = df02)
summary(lm01)
summary(lm02)

stepAIC(lm01, direction="both", trace=0,)
stepAIC(lm02, direction="both", trace=0,)
stepAIC(lm01, direction="both", trace=0, k = log(95))
stepAIC(lm02, direction="both", trace=0, k = log(95))
```
## SUR: Step AIC/BIC
```{r}
lm03 = lm(fm01, data = df03)
lm04 = lm(fm02, data = df03)
summary(lm03)
summary(lm04)

stepAIC(lm03, direction="both", trace=0,)
stepAIC(lm04, direction="both", trace=0,)
stepAIC(lm03, direction="both", trace=0, k = log(95))
stepAIC(lm04, direction="both", trace=0, k = log(95))
```

## SCR: Lasso
```{r}
par(mfrow=c(1,1))
lasso01 <- glmnet(ma02[,-1], ma02[,1])
plot(lasso01, label = T)
print("coefficients values:")
coef(lasso01, s = 0.02)
```

## SUR: Lasso
```{r}
par(mfrow=c(1,1))
lasso02 <- glmnet(ma03[,-1], ma03[,1])
plot(lasso02, label = T)
print("coefficients values:")
coef(lasso01, s = 0.02)
```

## SCR: Lasso with cross validation
```{r}
lasso.cv01 <- cv.glmnet(ma02[,-1], ma02[,1])
plot(lasso.cv01)

lasso.cv01$lambda.min
lasso.cv01$lambda.1se
coef(lasso.cv01, s = "lambda.min")
coef(lasso.cv01, s = "lambda.1se")
```

## SUR: Lasso with cross validation
```{r}
lasso.cv02 <- cv.glmnet(ma03[,-1], ma03[,1])
plot(lasso.cv02)

lasso.cv02$lambda.min
lasso.cv02$lambda.1se
coef(lasso.cv02, s = "lambda.min")
coef(lasso.cv02, s = "lambda.1se")
```

## SCR: boosting with linear base learner
```{r}
boost00 = glmboost(fm01, data = df02)
coef(boost00, which = "")
plot(boost00, ylim=range(coef(boost00, which = colnames(df02)[5:10])))

boost01 = gamboost(fm01, data = df02,
                   family=Gaussian(),
                   baselearner="bols",
                   boost_control(mstop = 5, nu=0.1))
coef(boost01, which = "")
par(mfrow=c(2,3))
#plot(boost01, off2int=TRUE)
plot(boost01)
cols_chosen = unique(boost01$xselect())
colnames(df02)[cols_chosen+4]
boost01$xselect()
#unique(boost01$xselect())

# With interactions
boost02 = glmboost(fm02, data = df02)
coef(boost02, which = "")
#plot(boost02, ylim=range(coef(boost02, which = "CCND1,RB1,CDKN1A,ESR1,CCNE1, MYC, CCND1:RB1, CCND1:CDKN1A, #CCND1:ESR1,CDKN1A:CCNE1, RB1:CCNE1, ESR1:MYC,CDKN1A:MYC" )))

boost03 = gamboost(fm02, 
                   data = df02,
                   baselearner="bols")
unique(boost03$xselect())
print("RB1, CCND1*RB1, CCND1, CCNE1, ESR1, MYC, CDKN1A")
```

## SCR: boosting with stump as base learner
```{r}
stump01 = gamboost(fm01, data = df02, 
                   family=Gaussian(), 
                   baselearner='btree',
                   boost_control(mstop = 10))
frame()
dev.new()
par(mfrow=c(1,2), mar=rep(4,4))
plot(stump01)
cols_chosen = unique(stump01$xselect())
colnames(df02)[cols_chosen+4]
stump01$xselect()
```

## Boosting with boostrap cross-validation
```{r}
boost00.cv = cvrisk(boost00)
par(mfrow=c(1,1))
plot(boost00.cv)
mstop(boost00.cv)
boost00[mstop(boost00.cv)]
names(coef(boost00))

boost01.cv = cvrisk(boost01)
plot(boost01.cv)
mstop(boost01.cv)
boost01[mstop(boost01.cv)]
names(coef(boost01))

stump01.cv = cvrisk(stump01)
plot(stump01.cv)
mstop(stump01.cv)
stump01[mstop(stump01.cv)]
names(coef(stump01))

```

# Graphs
```{r echo=FALSE}
cor.ma02 <- cor(ma02)

pc.fit.10prec <- pc(suffStat = list(C = cor.ma02, n = 100),
             indepTest = gaussCItest, # indep.test: partial correlations
             alpha=0.1, 
             labels = colnames(cor.ma02), 
             verbose = TRUE)
pc.fit.5prec <- pc(suffStat = list(C = cor.ma02, n = 100),
             indepTest = gaussCItest, # indep.test: partial correlations
             alpha=0.05, 
             labels = colnames(cor.ma02), 
             verbose = TRUE)
```

```{r}
par(mfrow=c(1,2))
plot(pc.fit.10prec, main = "Estimated CPDAG (alpha=10%)")
plot(pc.fit.5prec, main = "Estimated CPDAG (alpha=5%)")
#as(pc.fit.10prec, "amat")
```

# B: PREDICTION

## plot gene(SCR) vs Prolif(SUR)
```{r echo=FALSE}
par(mfrow=c(2,3), mar=rep(2,4))
names = c('CCND1', 'CCNE1', 'CDKN1A', 'ESR1', 'MYC','RB1') 
for(i in 1:6){
  x = ma04[,i+1]
  y = ma04[,1]
  plot(x, y, main=names[i])
  abline(lm(y ~ x), col = "red", lwd=3)
  mtext(paste("Correlation:", round(cor(x, y), 2)), side=1, line=-1.5, cex=1.2)
  cat(names[i])
  print(cor(x,y), method = c("pearson"))
}
```

## Lasso prediction
```{r}
X  <- subset(df05, select= -ProliferationScore)
Y  <- subset(df05, select= ProliferationScore)

prediction.lasso <- function(X, Y){
  X = as.matrix(X)
  Y = as.matrix(Y)
  sample = sample.int(n=48, size=45)
  X_train = X[sample,]
  X_test = X[-sample,]
  Y_train = Y[sample,]
  Y_test = Y[-sample,]

  lasso.cv <- cv.glmnet(X_train, Y_train)
  #lasso.pred = predict(lasso, newx = as.matrix(X_test), type = "response", s = "lambda.1se")
  lasso.pred = predict(lasso.cv, newx = X_test, type = "response", s = "lambda.min")
  out = data.frame(lasso.pred, Y_test)
  return(out)
}

# prediction.lasso(X,Y)
set.seed(111)

pred.lasso.loop = function(X, Y){
  df_ <- data.frame(matrix(ncol = 2, nrow = 0))
  colnames(df_) <- c("lambda.min", "Y_test")

  for (i in c(1:100)) {
    df_ = rbind(df_, prediction.lasso(X,Y))
  }  
  return(df_)
}

df06 = pred.lasso.loop(X,Y)
```
## Results: Lasso prediction
```{r}
plot(df06[,1], df06[,2])
abline(lm(df06[,2] ~ df06[,1]), col = "red", lwd=3)
mtext(paste("Correlation:", round(cor(df06[,2], df06[,1]), 3)), side=1, line=-1, cex=1.4)

mean((df06[,1] - df06[,2])^2)
```

## Stump prediction
```{r}
prediction.stump <- function(){
  sample = sample.int(n=48, size=45)
  df_train = df05[sample,]
  df_test = df05[-sample,]

  stump = gamboost(fm01, data = df_train, 
                   family=Gaussian(), 
                   baselearner='btree',
                   boost_control(mstop = 100))

  stump.cv = cvrisk(stump)
  stump.pred = predict(stump, newdata = df_test, type = "response")

  out = data.frame(stump.pred, df_test[,1])
  return(out)
}

prediction.stump()
set.seed(111)

pred.stump.loop = function(){
  df_ <- data.frame(matrix(ncol = 2, nrow = 0))
  colnames(df_) <- c("stump.pred", "ProliferationScore")

  for (i in c(1:100)) {
    df_ = rbind(df_, prediction.stump())
  }  
  return(df_)
}

df07 <- pred.stump.loop()
```

## Results: Stump prediction
```{r}
plot(df07[,1], df07[,2])
abline(lm(df07[,2] ~ df07[,1]), col = "red", lwd=3)
mtext(paste("Correlation:", round(cor(df07[,2], df07[,1]), 3)), side=1, line=-1, cex=1.4)

mean((df07[,1] - df07[,2])^2)
```

## Prediction only at SCR
```{r}

df08 = subset(df02, select=-c(UniqueID, timepoint, TrialArmNeo))

pred.lasso.stump = function(){
  sample = sample.int(n=95, size=90)
  df_train = df08[sample,]
  df_test = df08[-sample,]

  lasso <- cv.glmnet(as.matrix(df_train[,2:7]), as.matrix(df_train[,1]))

  # lasso.pred = predict(lasso, newx = as.matrix(X_test), s = "lambda.1se")
  lasso.pred = predict(lasso, newx = as.matrix(df_test[,2:7]), type = "response", s = "lambda.min")

  stump = gamboost(fm01, data = df_train, 
                   family=Gaussian(), 
                   baselearner='btree',
                   boost_control(mstop = 100))

  stump.cv = cvrisk(stump)
  stump.pred = predict(stump, newdata = df_test, type = "response")

  out = data.frame(lasso.pred, stump.pred, df_test[,1])
  return(out)
}

pred.lasso.stump()

pred.l.s = pred.lasso.stump()

set.seed(111)

pred.l.s.loop = function(){
  df_ <- data.frame(matrix(ncol = 3, nrow = 0))
  colnames(df_) <- c("lambda.min" ,"stump.pred", "ProliferationScore")

  for (i in c(1:10)) {
    df_ = rbind(df_,  pred.lasso.stump())
  }  
  return(df_)
}

df.l.s <- pred.l.s.loop()

par(mfrow=c(1,2), mar=rep(2,4))
plot(df.l.s$lambda.min, df.l.s$ProliferationScore, main="Lasso")
abline(lm(df.l.s[,3] ~ df.l.s[,1]), col = "red", lwd=3)
mtext(paste("Correlation:", round(cor(df.l.s[,3], df.l.s[,1]), 3)), side=1, line=-1, cex=1.4)

plot(df.l.s$stump.pred, df.l.s$ProliferationScore, , main="Stump")
abline(lm(df.l.s[,3] ~ df.l.s[,2]), col = "red", lwd=3)
mtext(paste("Correlation:", round(cor(df.l.s[,3], df.l.s[,2]), 3)), side=1, line=-1, cex=1.4)

mean((df.l.s$lambda.min - df.l.s$ProliferationScore)**2)
mean((df.l.s$stump.pred - df.l.s$ProliferationScore)**2)
```