### 2023-01-30 R_codeP01/reports/test01.Rmd
worked on report

---
title: "Milestone 2: Core results 01"
author: "Anders"
date: "25/1/2023"
output: pdf_document
#output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r}
report02_out <- function(object){

  fr <- sum(is.na(object$correlations))/length(object$correlations)
  cat("Fraction of model fits with no selected genes:", fr, "\n")
  cat("Correlations mean: ", mean(na.omit(lasso_k_ob$correlations)))

  # Histogram correlations
  correlations_finite <- lasso_k_ob$correlations[is.finite(lasso_k_ob$correlations)]
  cor_df <- data.frame(correlation = correlations_finite)

  ggplot(cor_df, aes(x=correlation)) +
    geom_histogram(bins = 30, color = "black", fill = "white") +
    xlab("Correlation") +
    ylab("Frequency") +
    ggtitle("Histogram of Correlation Values")
}
```

```{r}
path <- "/Users/anders/Documents/MASTER/Cancer/model_instances/lk_AllGenesProlif.RData"
load(path)

report02_out(lasso_k_ob)
```