### 2023-01-30 R_codeP01/reports/MS2_report02.Rmd
Worked on report

---
title: "Milestone 2: Core results 01"
author: "Anders"
date: "25/1/2023"
output: pdf_document
#output: html_notebook
editor_options: 
  chunk_output_type: console
---

<!-- ```{r echo=FALSE}  -->
<!-- knitr::opts_chunk$set(echo=FALSE)  -->
<!-- ``` -->

# Data
One clinical trails on breast cancer (advanced HR+/HER2- and HER2-E breast cancer) using two different drug combination; and a cohorts study (here used as test data set). /
Both data set have 771 genes. This genes are specifically selected based likely potential role in breast cancer. /
The gene set is dived into X numbers of "signature genes"; which are thought to represent unities with respect to cancer biology...

## Trail
Two treatments which differ with respect to drug combination 
- Target: ribociclib and endocrine therapy (letrozole)
- Chemotherapy: doxorubicin, cyclophosphamide and paclitaxel. 
approx. 50 patients in each group.
Endpoints: proliferation score, ROR score

## Cohort
The primary objective of this study is to compare two cdk4/6 targeted drugs (Palbociclib, n=36; Abemaciclib, n=3 in combination with endocrine therapy (tamoxifen, fulvestrant or aromatase inhibitors,  I think?)

Endpoints: progression free survival (months), OS?, and status of the two former (dont know what that means)

# Major goal
1. Find best model to predict outcome of cancer treatment with genetic profile as predictive features
2. Features selection in order to understand cancer biology

# Major challanges
Preliminary experiments (on trail 1) showed instability in prediction and feature selection between bootstrap samples of Lasso. I believe this is a classical problem of high-dim data?

# Approch
Test all thinkable models to see if some is surperior

# Evaluation of models
Two levels of evaluation is scheduled:

## 1. Relative comparison of models
1000 bootstrap models are fitted and then evaluated on the original sample. This gives a relative comparison of the various models with respect to data similar to the given data set. In addition to Correlation and MSE, frequency of selected features is compared.

## 2. Expected outcome of future patients
3 strategies are considered:  
1. Repeated cross-validations (100 rep) \
2. Bootstrap models with 0.632 (or 0.632?) adjustment \
3. Test data set from a second trail (This trail have different responses) \

# Models
## Lasso
## Post Lasso
## Ridge
## Elastic Net
## Boosting with stumps as base learner
### - mboost
### - xgboost
## Feature selecting ensemble model

# RESULTS
```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r}
source("/Users/anders/Documents/Master/Cancer/R_codeP01/scripts/90_output01.R")
```

# Lasso
```{r}
path <- "/Users/anders/Documents/MASTER/Cancer/model_instances/lk_AllGenesProlif.RData"
load(path)

results_lasso <- report02_out(lasso_k_ob)
```

# Ridge
# Elastic Net
# Boosting with stumps as base learner

# Post Lasso