### 2023-02-01 R_codeP01/reports/MS2_report02.Rmd
xgboost w/rep.cv; residuals w/lasso

---
title: "Milestone 2: Core results 01"
author: "Anders"
date: "25/1/2023"
output: pdf_document
#output: html_notebook
editor_options: 
  chunk_output_type: console
---

<!-- ```{r echo=FALSE}  -->
<!-- knitr::opts_chunk$set(echo=FALSE)  -->
<!-- ``` -->
```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r}
# lb_obj_6_prolif.RData
# lb_obj_771_prolif.RData
# lb_obj_771_residuals_prolif.RData
# lb_obj_nodes_prolif.RData

# lb_obj_771_ROR.RData
#
# lk_obj_771_prolif.RData
#
# bb_obj_6_prolif.RData
# 
#
```

# Data
One clinical trails on breast cancer (advanced HR+/HER2- and HER2-E breast cancer) using two different drug combination; and a cohorts study (here used as test data set). /
Both data set have mRNA expression of 771 genes at baseline (prior to treatment). This genes are specifically selected based on their potential roles in breast cancer pathology: /
The gene set is dived into X numbers of "signature genes"; which are thought to represent functional unities with respect to cancer biology...overview of signatures..

Additionally, both data-set contians clinical data as...
(how to model in...)

## Respones used
### Proliferation score
### Risk of relapse score (ROR)
### ...

## Trail
Two treatments which differ with respect to drug combination 
- Target: ribociclib and endocrine therapy (letrozole)
- Chemotherapy: doxorubicin, cyclophosphamide and paclitaxel. 
approx. 50 patients in each group.
Endpoints: proliferation score, ROR score, combined ROR and prolif

## Cohort
The primary objective of this study is to compare two cdk4/6 targeted drugs (Palbociclib, n=36; Abemaciclib, n=3 in combination with endocrine therapy (tamoxifen, fulvestrant or aromatase inhibitors,  I think?)

Endpoints: progression free survival (months), OS?, and status of the two former (dont know what that means)

# Major goal
1. Find best model to predict outcome of cancer treatment with genetic profile as predictive features
2. Features selection in order to understand cancer biology

# Major challanges
Preliminary experiments (on trail 1) showed instability in prediction and feature selection between bootstrap samples of Lasso. I believe this is a classical problem of high-dim data?

# Approch
Test all thinkable models to see if some is superior

# Evaluation of models
Two levels of evaluation is scheduled:

## 1. Relative comparison of models
1000 bootstrap models are fitted and then evaluated on the original sample. This gives a relative comparison of the various models with respect to data very similar to the given data set. In addition to Correlation and MSE, frequency of selected features is compared.

## 2. Expected outcome of future patients
3 strategies are considered:  
1. Repeated cross-validations (200 rep, 5-fold) \
2. Bootstrap models with 0.632 (or 0.632?) adjustment (Not done) \
3. Use the cohort as test data-set (Challenge: This trail have different responses) \

# Models
## Lasso
## Post Lasso
## Residuals
## Ridge
## Elastic Net
## Boosting with stumps as base learner
### - mboost
### - xgboost
## Feature selecting ensemble model

# RESULTS
```{r}
source("/Users/anders/Documents/Master/Cancer/R_codeP01/scripts/90_output01.R")
```

# Lasso

## Bootstrap
### 6 genes -> proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_6_prolif.RData")
l_b_6_p <- report02_out(lb_obj_6_prolif)
```
### 6 genes -> ROR_proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_6_RORprolif.RData")
l_b_6_ROR_p <- report02_out(lb_obj_6_RORprolif)
```

### 771 genes -> proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_771_prolif.RData")
l_b_771_p <- report02_out(lb_obj_771_prolif)
```
### 771 genes -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_771_RORprolif.RData")
l_b_771_ROR_p <- report02_out(lb_obj_771_RORprolif)
```

### node values -> proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_nodes_prolif.RData")
l_b_nodes_p <- report02_out(lb_obj_nodes_prolif)
```
### node values -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_nodes_RORprolif.RData")
l_b_nodes_ROR_p <- report02_out(lb_obj_nodes_RORprolif)
```

### Mechnaistic + Residuals -> proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_residuals771_prolif.RData")
l_b_residuals771_p <- report02_out(lb_obj_residuals771_prolif)
```
### Mechnaistic + Residuals -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_residuals_RORprolif.RData")
l_b_residuals_ROR_p <- report02_out(lb_obj_residuals_RORprolif)
```

## Repeated cross-validation
### 6 genes -> proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_6_prolif.RData")
l_c_6_p <- report02_out(lc_obj_6_prolif)
```
### 6 genes -> ROR_proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_6_RORprolif.RData")
l_c_6_ROR_p <- report02_out(lc_obj_6_RORprolif)
```

### 771 genes -> proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_771_prolif.RData")
l_c_771_p <- report02_out(lc_obj_771_prolif)
```
### 771 genes -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_771_RORprolif.RData")
l_c_771_ROR_p <- report02_out(lc_obj_771_RORprolif)
```

### node values -> proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_nodes_prolif.RData")
l_c_nodes_p <- report02_out(lc_obj_nodes_prolif)
```
### node values -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_nodes_RORprolif.RData")
l_c_nodes_ROR_p <- report02_out(lc_obj_nodes_RORprolif)
```

# Ridge
# Elastic Net
# Boosting with stumps as base learner

# Post Lasso