### 2023-02-11 R_codeP01/reports/MS2_report02.Rmd
elastic net boot and rc

---
title: "Milestone 2: Core results 01"
author: "Anders"
date: "25/1/2023"
output: pdf_document
#output: html_notebook
editor_options: 
  chunk_output_type: console
---

<!-- ```{r echo=FALSE}  -->
<!-- knitr::opts_chunk$set(echo=FALSE)  -->
<!-- ``` -->
```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r}
## INPUT (features):
# genes, demographics, clinical

## OUTPUT (response):
# proliferation score
# ROR score
# ROR: low, mid, high
# ROR+prolif score
# ROR+prolif: low, mid, high
# others? e.g.: Cellularity_perc, Ki67_perc, Positive_lymph_nodes ...

## OBJECTS:
# lb_obj_6_prolif.RData
# lb_obj_771_prolif.RData
# lb_obj_residuals_6_prolif.RData
# lb_obj_771_residuals_prolif.RData
# lb_obj_nodes_prolif.RData

# lb_obj_6_ROR_prolif.RData
# lb_obj_771_ROR_prolif.RData
# lb_obj_771_residuals_ROR_prolif.RData
# similar objects with logistic regression on low vs (mid+high)

# lc_obj_6_prolif.RData
# lc_obj_771_prolif.RData
# lc_obj_771_residuals_prolif.RData
# lc_obj_nodes_prolif.RData
#
# lc_obj_6_ROR_prolif.RData
# lc_obj_771_ROR_prolif.RData
# lc_obj_771_residuals_ROR_prolif.RData
# similar objects with logistic regression on low vs (mid+high)
# ...
# rb_obj_6_prolif.RData
#
# eb_obj_6_prolif.RData
#
# bb_obj_6_prolif.RData
# 
# bc_obj_6_prolif.RData
#
# pb_obj_6_prolif.RData
#
# sb_obj_6_prolif.RData
#
# cb_obj_6_prolif.RData

```

# Data
One clinical trails on breast cancer (advanced HR+/HER2- and HER2-E breast cancer) using two different drug combination; and a cohorts study (here used as test data set).  
Both data set have mRNA expression of 771 genes at baseline (prior to treatment). This genes are specifically selected based on their potential roles in breast cancer pathology:  
The gene set is dived into 25 sets of "signature genes"; which are thought to represent functional unities with respect to cancer biology. Often signaling pathways. Furthermore, 8 immune cells are represented with specific genes. These sets are substantially smaller then the signature genes; which I presume leads to some issue in modeling (as for clinical data too - see next sentence).

Additionally, both data-set contains clinical data; which up to now is not used in any models. If included they maybe should have a higher weight or be implemented differently from a sole gene. Maybe in a stacked ensemble model as signature.

## Respones used
### Proliferation score
A score based on expression level of some of the genes. Range: -1.1366 to 0.8511
### Risk of relapse score (ROR)
A score based on expression level of some of the genes. Range: -8.035678 to 75.13174
### Risk of relapse score with proliferaton score (ROR-Prolif)
A combined score based on expression level of more genes. Range: 1 to 97 (1-100).

The two scores involving ROR also have categorical variants containing: \ 
low, medium, high

### Progression free survival (PFS)
This is the outcome used in the clinical cohort. Here i have used correlation with the scores described above. Spearman can maybe be used. Another approach is to use the above scores to divide the patient in to two groups and see if the two groups show clearly separable PFS over time (basically lock at the graphs). The differentiation into two groups is done base on best values from a ROC curve.

## Trail
Two treatments which differ with respect to drug combination 
- Target: ribociclib and endocrine therapy (letrozole)
- Chemotherapy: doxorubicin, cyclophosphamide and paclitaxel. 
approx. 50 patients in each group.
Endpoints: proliferation score, ROR score, combined ROR and prolif

## Cohort
The primary objective of this study is to compare two cdk4/6 targeted drugs (Palbociclib, n=36; Abemaciclib, n=3 in combination with endocrine therapy (tamoxifen, fulvestrant or aromatase inhibitors,  I think?)

Endpoints: progression free survival (months), OS?, and status of the two former (dont know what that means)

# Major goal
1. Find best model to predict outcome of cancer treatment with genetic profile as predictive features
2. Features selection in order to understand cancer biology

# Major challanges
Preliminary experiments (on trail 1) showed instability in prediction and feature selection between bootstrap samples of Lasso. I believe this is a classical problem of high-dim data?

# Approch
Test all thinkable models in a search for superior models

# Evaluation of models
Two levels of evaluation is considered:

## 1. Relative comparison of the different models
1000 bootstrap models are fitted and then evaluated on the original sample. This gives a relative comparison of the various models with respect to data very similar to the given data set. In addition to Correlation and MSE, frequency of selected features is compared.

## 2. Expected outcome of future patients
3 strategies are considered:  
1. Repeated cross-validations (200 rep, 5-fold) \ 
2. Bootstrap models with 0.632 (or 0.632?) adjustment (Not done) \ 
3. Use the cohort as test data-set (Challenge: This trail have different responses) \

# RESULTS

# Results of individual modles:

# Models tested
## Lasso
## Post Lasso
## Residuals
## Ridge
## Elastic Net
## Boosting with stumps as base learner
### - mboost
### - xgboost
## PCA feature engineering
## Stacking using different features in the base models 
## Synergistic learning
```{r}
source("/Users/anders/Documents/Master/Cancer/R_codeP01/scripts/90_output01.R")
source("/Users/anders/Documents/Master/Cancer/R_codeP01/scripts/90_output_ridge_01.R")
```

# Lasso - Bootstrap
### 6 genes -> proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_6_prolif.RData")
l_b_6_p <- report02_out(lb_obj_6_prolif)
```
### 6 genes -> ROR_proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_6_RORprolif.RData")
l_b_6_ROR_p <- report02_out(lb_obj_6_RORprolif)
```

### 771 genes -> proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_771_prolif.RData")
l_b_771_p <- report02_out(lb_obj_771_prolif)
```
### 771 genes -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_771_RORprolif.RData")
l_b_771_ROR_p <- report02_out(lb_obj_771_RORprolif)
```

### node values -> proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_nodes_prolif.RData")
l_b_nodes_p <- report02_out(lb_obj_nodes_prolif)
```
### node values -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_nodes_RORprolif.RData")
l_b_nodes_ROR_p <- report02_out(lb_obj_nodes_RORprolif)
```

### Mechanistic + Residuals -> proliferation score (additive)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_residuals_771_prolif_a.RData")
l_b_residuals_771_p_a <- report02_out(lb_obj_residuals_771_prolif_a)
```
### Mechanistic + Residuals -> proliferation score (multiplicative)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_residuals_771_prolif_m.RData")
l_b_residuals_771_p_m <- report02_out(lb_obj_residuals_771_prolif_m)
```
### Mechnaistic + Residuals -> ROR-proliferation score (additive)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_residuals_771_RORprolif_a.RData")
l_b_residuals_ROR_p_a <- report02_out(lb_obj_residuals_771_RORprolif_a)
```
### Mechnaistic + Residuals -> ROR-proliferation score (multiplicative)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_residuals_771_RORprolif_m.RData")
l_b_residuals_ROR_p_m <- report02_out(lb_obj_residuals_771_RORprolif_m)
```

### Summery results: lasso proliferation score (bootstrap) 
```{r}
library(knitr)
data_table <- data.frame(
  Model = c("lasso 6 genes", "lasso 771 genes", "Nodes", "Residual additive", "Residual multiplicative"),
  cor_mean = c(l_b_6_p$cor_mean, l_b_771_p$cor_mean, l_b_nodes_p$cor_mean, l_b_residuals_771_p_a$cor_mean, l_b_residuals_771_p_m$cor_mean),
  sd_cor = c(l_b_6_p$cor_sd, l_b_771_p$cor_sd, l_b_nodes_p$cor_sd, l_b_residuals_771_p_a$cor_sd, l_b_residuals_771_p_m$cor_sd),
  MSE_mean = c(l_b_6_p$MSE_mean, l_b_771_p$MSE_mean, l_b_nodes_p$MSE_mean, l_b_residuals_771_p_a$MSE_mean, l_b_residuals_771_p_m$MSE_mean),
  MSE_sd = c(l_b_6_p$MSE_sd, l_b_771_p$MSE_sd, l_b_nodes_p$MSE_sd, l_b_residuals_771_p_a$MSE_sd, l_b_residuals_771_p_m$MSE_sd)
)
kable(data_table)
```

### Summery results: lasso ROR+proliferation score (bootstrap) 
```{r}
data_table <- data.frame(
  Model = c("lasso 6 genes", "lasso 771 genes", "Nodes", "Residual additive", "Residual multiplicative"),
  cor_mean = c(l_b_6_ROR_p$cor_mean, l_b_771_ROR_p$cor_mean, l_b_nodes_ROR_p$cor_mean, l_b_residuals_ROR_p_a$cor_mean, l_b_residuals_ROR_p_m$cor_mean),
  sd_cor = c(l_b_6_ROR_p$cor_sd, l_b_771_ROR_p$cor_sd, l_b_nodes_ROR_p$cor_sd, l_b_residuals_ROR_p_a$cor_sd, l_b_residuals_ROR_p_m$cor_sd),
  MSE_mean = c(l_b_6_ROR_p$MSE_mean, l_b_771_ROR_p$MSE_mean, l_b_nodes_ROR_p$MSE_mean, l_b_residuals_ROR_p_a$MSE_mean, l_b_residuals_ROR_p_m$MSE_mean),
  MSE_sd = c(l_b_6_ROR_p$MSE_sd, l_b_771_ROR_p$MSE_sd, l_b_nodes_ROR_p$MSE_sd, l_b_residuals_ROR_p_a$MSE_sd, l_b_residuals_ROR_p_m$MSE_sd)
)
kable(data_table)
```

## Lasso - Repeated cross-validation
200 repeats of five fold cross-validation \ 
### 6 genes -> proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_6_prolif.RData")
l_c_6_p <- report02_out(lc_obj_6_prolif)
```
### 6 genes -> ROR_proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_6_RORprolif.RData")
l_c_6_ROR_p <- report02_out(lc_obj_6_RORprolif)
```

### 771 genes -> proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_771_prolif.RData")
l_c_771_p <- report02_out(lc_obj_771_prolif)
```
### 771 genes -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_771_RORprolif.RData")
l_c_771_ROR_p <- report02_out(lc_obj_771_RORprolif)
```

### node values -> proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_nodes_prolif.RData")
l_c_nodes_p <- report02_out(lc_obj_nodes_prolif)
```
### node values -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_nodes_RORprolif.RData")
l_c_nodes_ROR_p <- report02_out(lc_obj_nodes_RORprolif)
```

### Mechanistic + Residuals -> proliferation score (additive)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_residuals_771_p_a.RData")
l_c_residuals_771_p_a <- report02_out(lc_obj_residuals_771_p_a)
```
### Mechanistic + Residuals -> proliferation score (multiplicative)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_residuals_771_p_m.RData")
l_c_residuals_771_p_m <- report02_out(lc_obj_residuals_771_p_m)
```
### Mechnaistic + Residuals -> ROR-proliferation score (additive)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_residuals_771_ROR_p_a.RData")
l_c_residuals_ROR_p_a <- report02_out(lc_obj_residuals_771_ROR_p_a)
```
### Mechnaistic + Residuals -> ROR-proliferation score (multiplicative)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_residuals_771_ROR_p_m.RData")
l_c_residuals_ROR_p_m <- report02_out(lc_obj_residuals_771_ROR_p_m)
```

### Summery results: lasso proliferation score (repeated cross-validation) 
```{r}
library(knitr)
data_table <- data.frame(
  Model = c("lasso 6 genes", "lasso 771 genes", "Nodes", "Residual additive", "Residual multiplicative"),
  cor_mean = c(l_c_6_p$cor_mean, l_c_771_p$cor_mean, l_c_nodes_p$cor_mean, l_c_residuals_771_p_a$cor_mean, l_c_residuals_771_p_m$cor_mean),
  sd_cor = c(l_c_6_p$cor_sd, l_c_771_p$cor_sd, l_c_nodes_p$cor_sd, l_c_residuals_771_p_a$cor_sd, l_c_residuals_771_p_m$cor_sd),
  MSE_mean = c(l_c_6_p$MSE_mean, l_b_771_p$MSE_mean, l_c_nodes_p$MSE_mean, l_c_residuals_771_p_a$MSE_mean, l_c_residuals_771_p_m$MSE_mean),
  MSE_sd = c(l_c_6_p$MSE_sd, l_c_771_p$MSE_sd, l_c_nodes_p$MSE_sd, l_c_residuals_771_p_a$MSE_sd, l_c_residuals_771_p_m$MSE_sd)
)
kable(data_table)
```

### Summery results: lasso ROR+proliferation score (repeated cross-validation) 
```{r}
data_table <- data.frame(
  Model = c("lasso 6 genes", "lasso 771 genes", "Nodes", "Residual additive", "Residual multiplicative"),
  cor_mean = c(l_c_6_ROR_p$cor_mean, l_c_771_ROR_p$cor_mean, l_c_nodes_ROR_p$cor_mean, l_c_residuals_ROR_p_a$cor_mean, l_c_residuals_ROR_p_m$cor_mean),
  sd_cor = c(l_c_6_ROR_p$cor_sd, l_c_771_ROR_p$cor_sd, l_c_nodes_ROR_p$cor_sd, l_c_residuals_ROR_p_a$cor_sd, l_c_residuals_ROR_p_m$cor_sd),
  MSE_mean = c(l_c_6_ROR_p$MSE_mean, l_c_771_ROR_p$MSE_mean, l_c_nodes_ROR_p$MSE_mean, l_c_residuals_ROR_p_a$MSE_mean, l_c_residuals_ROR_p_m$MSE_mean),
  MSE_sd = c(l_c_6_ROR_p$MSE_sd, l_c_771_ROR_p$MSE_sd, l_c_nodes_ROR_p$MSE_sd, l_c_residuals_ROR_p_a$MSE_sd, l_c_residuals_ROR_p_m$MSE_sd)
)
kable(data_table)
```

# Ridge bootstrap
### 771 genes -> proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/r_b_obj_771_p.RData")
r_b_771_p <- report02_out_ridge(r_b_obj_771_p)
```
### 771 genes -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/r_b_obj_771_ROR_p.RData")
r_b_771_ROR_p <- report02_out_ridge(r_b_obj_771_ROR_p)
```

# Ridge repeated cross-validation
### 771 genes -> proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/r_c_obj_771_prolif.RData")
r_c_771_p <- report02_out_ridge(r_c_obj_771_prolif)
```
### 771 genes -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/r_c_obj_771_RORprolif.RData")
r_c_771_ROR_p <- report02_out_ridge(r_c_obj_771_RORprolif)
```

### Summery results: ridge 771 genes bootstrap and repeated cross-validation
```{r}
data_table <- data.frame(
  Model = c("prolif boot", "ROR-prolif boot", "prolif rep cross-val", "ROR-prolif rep cross-val"),
  cor_mean = c(r_b_771_p$cor_mean, r_b_771_ROR_p$cor_mean, r_c_771_p$cor_mean, r_c_771_ROR_p$cor_mean),
  sd_cor = c(r_b_771_p$cor_sd, r_b_771_ROR_p$cor_sd, r_c_771_p$cor_sd, r_c_771_ROR_p$cor_sd),
  MSE_mean = c(r_b_771_p$MSE_mean, r_b_771_ROR_p$MSE_mean, r_c_771_p$MSE_mean, r_c_771_ROR_p$MSE_mean),
  MSE_sd = c(r_b_771_p$MSE_sd, r_b_771_ROR_p$MSE_sd, r_c_771_p$MSE_sd, r_c_771_ROR_p$MSE_sd)
)
kable(data_table)
```

# Elastic Net - bootstrap
### 771 genes -> proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/e_b_obj_771_p.RData")
e_b_771_p <- report02_out_ridge(e_b_obj_771_p)
```
### 771 genes -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/e_b_obj_771_ROR_p.RData")
e_b_771_ROR_p <- report02_out_ridge(e_b_obj_771_ROR_p)
```

# Elastic Net: cross-validation
### 771 genes -> proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/e_c_obj_771_prolif.RData")
e_c_771_p <- report02_out_ridge(e_c_obj_771_prolif)
```
### 771 genes -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/e_c_obj_771_RORprolif.RData")
e_c_771_ROR_p <- report02_out_ridge(e_c_obj_771_RORprolif)
```

### Summery results: elastic net 771 genes bootstrap and repeated cross-validation
```{r}
data_table <- data.frame(
  Model = c("prolif boot", "ROR-prolif boot", "prolif rep cross-val", "ROR-prolif rep cross-val"),
  cor_mean = c(e_b_771_p$cor_mean, e_b_771_ROR_p$cor_mean, e_c_771_p$cor_mean, e_c_771_ROR_p$cor_mean),
  sd_cor = c(e_b_771_p$cor_sd, e_b_771_ROR_p$cor_sd, e_c_771_p$cor_sd, e_c_771_ROR_p$cor_sd),
  MSE_mean = c(e_b_771_p$MSE_mean, e_b_771_ROR_p$MSE_mean, e_c_771_p$MSE_mean, e_c_771_ROR_p$MSE_mean),
  MSE_sd = c(e_b_771_p$MSE_sd, e_b_771_ROR_p$MSE_sd, e_c_771_p$MSE_sd, e_c_771_ROR_p$MSE_sd)
)
kable(data_table)
```

# Boosting with stumps as base learner

# Post Lasso

# Summery of results
## Lasso proliferation score (bootstrap) 
```{r}
data_table <- data.frame(
  Model = c("lasso 6 genes", "lasso 771 genes", "Nodes", "Residual additive", "Residual multiplicative"),
  cor_mean = c(l_b_6_p$cor_mean, l_b_771_p$cor_mean, l_b_nodes_p$cor_mean, l_b_residuals_771_p_a$cor_mean, l_b_residuals_771_p_m$cor_mean),
  sd_cor = c(l_b_6_p$cor_sd, l_b_771_p$cor_sd, l_b_nodes_p$cor_sd, l_b_residuals_771_p_a$cor_sd, l_b_residuals_771_p_m$cor_sd),
  MSE_mean = c(l_b_6_p$MSE_mean, l_b_771_p$MSE_mean, l_b_nodes_p$MSE_mean, l_b_residuals_771_p_a$MSE_mean, l_b_residuals_771_p_m$MSE_mean),
  MSE_sd = c(l_b_6_p$MSE_sd, l_b_771_p$MSE_sd, l_b_nodes_p$MSE_sd, l_b_residuals_771_p_a$MSE_sd, l_b_residuals_771_p_m$MSE_sd)
)
kable(data_table)
```