### 2023-03-26 R_codeP01/reports/MS2_report03_truncated.Rmd
report writing

---
title: "Milestone 2: Core results 01"
author: "Anders"
date: "25/1/2023"
output: pdf_document
#output: html_notebook
editor_options: 
  chunk_output_type: console
---

<!-- ```{r echo=FALSE}  -->
<!-- knitr::opts_chunk$set(echo=FALSE)  -->
<!-- ``` -->
```{r global-options, include=FALSE}
library(tidyverse)
library(knitr)
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r}
## INPUT (features):
# genes, demographics, clinical

## OUTPUT (response):
# proliferation score
# ROR score
# ROR: low, mid, high
# ROR+prolif score
# ROR+prolif: low, mid, high
# others? e.g.: Cellularity_perc, Ki67_perc, Positive_lymph_nodes ...

## OBJECTS:
# lb_obj_6_prolif.RData
# lb_obj_771_prolif.RData
# lb_obj_residuals_6_prolif.RData
# lb_obj_771_residuals_prolif.RData
# lb_obj_nodes_prolif.RData

# lb_obj_6_ROR_prolif.RData
# lb_obj_771_ROR_prolif.RData
# lb_obj_771_residuals_ROR_prolif.RData
# similar objects with logistic regression on low vs (mid+high)

# lc_obj_6_prolif.RData
# lc_obj_771_prolif.RData
# lc_obj_771_residuals_prolif.RData
# lc_obj_nodes_prolif.RData
#
# lc_obj_6_ROR_prolif.RData
# lc_obj_771_ROR_prolif.RData
# lc_obj_771_residuals_ROR_prolif.RData
# similar objects with logistic regression on low vs (mid+high)
# ...
# rb_obj_6_prolif.RData
#
# eb_obj_6_prolif.RData
#
# bb_obj_6_prolif.RData
# 
# bc_obj_6_prolif.RData
#
# pb_obj_6_prolif.RData
#
# sb_obj_6_prolif.RData
#
# cb_obj_6_prolif.RData

```

# Data
One clinical trails on breast cancer (advanced HR+/HER2-) using two different drug combination

The data set have mRNA expression of 771 genes at baseline (prior to treatment). This genes are specifically selected based on their potential roles in breast cancer pathology:

The gene set is dived into 25 sets of "signature genes"; which are thought to represent functional unities with respect to cancer biology. Often signaling pathways. Furthermore, 8 immune cells are represented with specific genes. These sets are substantially smaller then the signature genes; which I presume leads to some issue in modeling (as for clinical data too - see next sentence).
In the domain knowledge part at the end 5 signature gene sets are use. Make sense to add more...

Additionally, the data-set contains clinical data; which up to now is not used in any models. If included they maybe should have a higher weight or be implemented differently from a sole gene. Maybe in a stacked ensemble model as signature.

## Respones in study -ish
### Proliferation score

A score based on expression level of some of the genes. Range: -1.1366 to 0.8511

### Risk of relapse score (ROR)

A combined score based on expression level of genes and some clinical findings. Range: -8.035678 to 75.13174

(only used in combination with proliferation score as described bellow)

### Risk of relapse score with proliferaton score (ROR_Prolif)

A combined score of the two above. Range: 1 to 97 (1-100)

The two scores involving ROR also have categorical variants containing: low, medium, high
(but not used...)

## Trail
Two treatments which differ with respect to drug combination 
- Target: ribociclib and endocrine therapy (letrozole)
- Chemotherapy: doxorubicin, cyclophosphamide and paclitaxel. 
approx. 50 patients in each group.
Endpoints: proliferation score, ROR score, combined ROR and prolif

# Major goal
1. Find best model to predict outcome of cancer treatment with genetic profile as predictive features
2. Features selection in order to understand cancer biology

# Major challanges
Preliminary experiments (on trail 1) showed instability in prediction and feature selection between bootstrap samples of Lasso. I believe this is a classical problem of high-dim data?

# Approch
Test all thinkable models in a search for superior models

# Evaluation of models
Two levels of evaluation is considered:

## 1. Relative comparison of the different models
1000 bootstrap models are fitted and then evaluated on the original sample. This gives a relative comparison of the various models with respect to data very similar to the given data set. Correlation, MSE and frequency of selected features is compared.

## 2. Expected outcome of future patients
3 strategies are considered:

1. Repeated cross-validations (200 rep, 5-fold)

2. Bootstrap models with 0.632 (or 0.632?) adjustment (Not done)

3. Use the cohort as test data-set (Challenge: This trail have different responses)

# RESULTS
## Features tested:
### 6 genes
### 771 genes
### node values of mech model
### residuals of mech model

## Responses tested:
### proliferation score
### combined score including proliferation and ROR

## Models tested:
### Lasso
### Ridge
### Elastic Net
### Boosting with stumps as base learner
#### (- mboost)
#### - xgboost
### PCA on subsets of genes
### Stacking using different features in the base models
### Sparse group lasso (not done)
### Iterative learning (ongoing)
### Post Lasso (not done)

# Results of individual modles:

```{r}
source("/Users/anders/Documents/Master/Cancer/R_codeP01/scripts/90_output01.R")
source("/Users/anders/Documents/Master/Cancer/R_codeP01/scripts/90_output_ridge_01.R")
```

# Lasso - Bootstrap
### 6 genes -> proliferation score (lasso - bootstrap)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_6_prolif.RData")
l_b_6_p <- report02_out(lb_obj_6_prolif)
```

### 6 genes -> ROR_proliferation score (lasso - bootstrap)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_6_RORprolif.RData")
l_b_6_ROR_p <- report02_out(lb_obj_6_RORprolif)
```

### 771 genes -> proliferation score (lasso - bootstrap)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_771_prolif.RData")
l_b_771_p <- report02_out(lb_obj_771_prolif)
```

### 771 genes -> ROR-proliferation score (lasso - bootstrap)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_771_RORprolif.RData")
l_b_771_ROR_p <- report02_out(lb_obj_771_RORprolif)
```

### node values -> proliferation score (lasso - bootstrap)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_nodes_prolif.RData")
l_b_nodes_p <- report02_out(lb_obj_nodes_prolif)
```

### node values -> ROR-proliferation score (lasso - bootstrap) 
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_nodes_RORprolif.RData")
l_b_nodes_ROR_p <- report02_out(lb_obj_nodes_RORprolif)
```

### Mechanistic + Residuals (additive) -> proliferation score  (lasso - bootstrap)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_residuals_771_prolif_a.RData")
l_b_residuals_771_p_a <- report02_out(lb_obj_residuals_771_prolif_a)
```

### Mechanistic + Residuals (multiplicative) -> proliferation score   (lasso - bootstrap)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_residuals_771_prolif_m.RData")
l_b_residuals_771_p_m <- report02_out(lb_obj_residuals_771_prolif_m)
```

### Mechnaistic + Residuals (additive) -> ROR-proliferation score  (lasso - bootstrap)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_residuals_771_RORprolif_a.RData")
l_b_residuals_ROR_p_a <- report02_out(lb_obj_residuals_771_RORprolif_a)
```

### Mechnaistic + Residuals (multiplicative) -> ROR-proliferation score (lasso - bootstrap)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lb_obj_residuals_771_RORprolif_m.RData")
l_b_residuals_ROR_p_m <- report02_out(lb_obj_residuals_771_RORprolif_m)
```

### Summery results: lasso proliferation score (bootstrap) 
```{r}
library(knitr)
data_table <- data.frame(
  Model = c("lasso 6 genes", "lasso 771 genes", "Nodes", "Residual additive", "Residual multiplicative"),
  cor_mean = c(l_b_6_p$cor_mean, l_b_771_p$cor_mean, l_b_nodes_p$cor_mean, l_b_residuals_771_p_a$cor_mean, l_b_residuals_771_p_m$cor_mean),
  sd_cor = c(l_b_6_p$cor_sd, l_b_771_p$cor_sd, l_b_nodes_p$cor_sd, l_b_residuals_771_p_a$cor_sd, l_b_residuals_771_p_m$cor_sd),
  MSE_mean = c(l_b_6_p$MSE_mean, l_b_771_p$MSE_mean, l_b_nodes_p$MSE_mean, l_b_residuals_771_p_a$MSE_mean, l_b_residuals_771_p_m$MSE_mean),
  MSE_sd = c(l_b_6_p$MSE_sd, l_b_771_p$MSE_sd, l_b_nodes_p$MSE_sd, l_b_residuals_771_p_a$MSE_sd, l_b_residuals_771_p_m$MSE_sd)
)
kable(data_table)
```

### Summery results: lasso ROR+proliferation score (bootstrap) 
```{r}
data_table <- data.frame(
  Model = c("lasso 6 genes", "lasso 771 genes", "Nodes", "Residual additive", "Residual multiplicative"),
  cor_mean = c(l_b_6_ROR_p$cor_mean, l_b_771_ROR_p$cor_mean, l_b_nodes_ROR_p$cor_mean, l_b_residuals_ROR_p_a$cor_mean, l_b_residuals_ROR_p_m$cor_mean),
  sd_cor = c(l_b_6_ROR_p$cor_sd, l_b_771_ROR_p$cor_sd, l_b_nodes_ROR_p$cor_sd, l_b_residuals_ROR_p_a$cor_sd, l_b_residuals_ROR_p_m$cor_sd),
  MSE_mean = c(l_b_6_ROR_p$MSE_mean, l_b_771_ROR_p$MSE_mean, l_b_nodes_ROR_p$MSE_mean, l_b_residuals_ROR_p_a$MSE_mean, l_b_residuals_ROR_p_m$MSE_mean),
  MSE_sd = c(l_b_6_ROR_p$MSE_sd, l_b_771_ROR_p$MSE_sd, l_b_nodes_ROR_p$MSE_sd, l_b_residuals_ROR_p_a$MSE_sd, l_b_residuals_ROR_p_m$MSE_sd)
)
kable(data_table)
```

## Lasso - Repeated cross-validation
200 repeats of five fold cross-validation

### 6 genes -> proliferation score (lasso - repeated cross-validation)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_6_prolif.RData")
l_c_6_p <- report02_out(lc_obj_6_prolif)
```
### 6 genes -> ROR_proliferation score (lasso - repeated cross-validation)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_6_RORprolif.RData")
l_c_6_ROR_p <- report02_out(lc_obj_6_RORprolif)
```

### 771 genes -> proliferation score (lasso - repeated cross-validation)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_771_prolif.RData")
l_c_771_p <- report02_out(lc_obj_771_prolif)
```

### 771 genes -> ROR-proliferation score (lasso - repeated cross-validation)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_771_RORprolif.RData")
l_c_771_ROR_p <- report02_out(lc_obj_771_RORprolif)
```

### node values -> proliferation score (lasso - repeated cross-validation)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_nodes_prolif.RData")
l_c_nodes_p <- report02_out(lc_obj_nodes_prolif)
```

### node values -> ROR-proliferation score (lasso - repeated cross-validation)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_nodes_RORprolif.RData")
l_c_nodes_ROR_p <- report02_out(lc_obj_nodes_RORprolif)
```

### Mechanistic + Residuals (additive) -> proliferation score (lasso - repeated cross-validation)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_residuals_771_p_a.RData")
l_c_residuals_771_p_a <- report02_out(lc_obj_residuals_771_p_a)
```

### Mechanistic + Residuals (multiplicative) -> proliferation score (lasso - repeated cross-validation)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_residuals_771_p_m.RData")
l_c_residuals_771_p_m <- report02_out(lc_obj_residuals_771_p_m)
```

### Mechnaistic + Residuals (additive) -> ROR-proliferation score (lasso - repeated cross-validation)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_residuals_771_ROR_p_a.RData")
l_c_residuals_ROR_p_a <- report02_out(lc_obj_residuals_771_ROR_p_a)
```

### Mechnaistic + Residuals (multiplicative) -> ROR-proliferation score (lasso - repeated cross-validation)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/lc_obj_residuals_771_ROR_p_m.RData")
l_c_residuals_ROR_p_m <- report02_out(lc_obj_residuals_771_ROR_p_m)
```

### Summery results: lasso proliferation score (repeated cross-validation) 
```{r}
library(knitr)
data_table <- data.frame(
  Model = c("lasso 6 genes", "lasso 771 genes", "Nodes", "Residual additive", "Residual multiplicative"),
  cor_mean = c(l_c_6_p$cor_mean, l_c_771_p$cor_mean, l_c_nodes_p$cor_mean, l_c_residuals_771_p_a$cor_mean, l_c_residuals_771_p_m$cor_mean),
  sd_cor = c(l_c_6_p$cor_sd, l_c_771_p$cor_sd, l_c_nodes_p$cor_sd, l_c_residuals_771_p_a$cor_sd, l_c_residuals_771_p_m$cor_sd),
  MSE_mean = c(l_c_6_p$MSE_mean, l_b_771_p$MSE_mean, l_c_nodes_p$MSE_mean, l_c_residuals_771_p_a$MSE_mean, l_c_residuals_771_p_m$MSE_mean),
  MSE_sd = c(l_c_6_p$MSE_sd, l_c_771_p$MSE_sd, l_c_nodes_p$MSE_sd, l_c_residuals_771_p_a$MSE_sd, l_c_residuals_771_p_m$MSE_sd)
)
kable(data_table)
```

### Summery results: lasso ROR+proliferation score (repeated cross-validation) 
```{r}
data_table <- data.frame(
  Model = c("lasso 6 genes", "lasso 771 genes", "Nodes", "Residual additive", "Residual multiplicative"),
  cor_mean = c(l_c_6_ROR_p$cor_mean, l_c_771_ROR_p$cor_mean, l_c_nodes_ROR_p$cor_mean, l_c_residuals_ROR_p_a$cor_mean, l_c_residuals_ROR_p_m$cor_mean),
  sd_cor = c(l_c_6_ROR_p$cor_sd, l_c_771_ROR_p$cor_sd, l_c_nodes_ROR_p$cor_sd, l_c_residuals_ROR_p_a$cor_sd, l_c_residuals_ROR_p_m$cor_sd),
  MSE_mean = c(l_c_6_ROR_p$MSE_mean, l_c_771_ROR_p$MSE_mean, l_c_nodes_ROR_p$MSE_mean, l_c_residuals_ROR_p_a$MSE_mean, l_c_residuals_ROR_p_m$MSE_mean),
  MSE_sd = c(l_c_6_ROR_p$MSE_sd, l_c_771_ROR_p$MSE_sd, l_c_nodes_ROR_p$MSE_sd, l_c_residuals_ROR_p_a$MSE_sd, l_c_residuals_ROR_p_m$MSE_sd)
)
kable(data_table)
```

# Ridge bootstrap
### 771 genes -> proliferation score (ridge bootstrap)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/r_b_obj_771_p.RData")
r_b_771_p <- report02_out_ridge(r_b_obj_771_p)
```

### 771 genes -> ROR-proliferation score (ridge bootstrap)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/r_b_obj_771_ROR_p.RData")
r_b_771_ROR_p <- report02_out_ridge(r_b_obj_771_ROR_p)
```

# Ridge repeated cross-validation
### 771 genes -> proliferation score (ridge repeated cross-validation)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/r_c_obj_771_prolif.RData")
r_c_771_p <- report02_out_ridge(r_c_obj_771_prolif)
```

### 771 genes -> ROR-proliferation score (ridge repeated cross-validation)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/r_c_771_RORprolif.RData")
r_c_771_ROR_p <- report02_out_ridge(r_c_771_RORprolif)
```

### Summery results: ridge 771 genes bootstrap and repeated cross-validation
```{r}
data_table <- data.frame(
  Model = c("prolif boot", "ROR-prolif boot", "prolif rep cross-val", "ROR-prolif rep cross-val"),
  cor_mean = c(r_b_771_p$cor_mean, r_b_771_ROR_p$cor_mean, r_c_771_p$cor_mean, r_c_771_ROR_p$cor_mean),
  sd_cor = c(r_b_771_p$cor_sd, r_b_771_ROR_p$cor_sd, r_c_771_p$cor_sd, r_c_771_ROR_p$cor_sd),
  MSE_mean = c(r_b_771_p$MSE_mean, r_b_771_ROR_p$MSE_mean, r_c_771_p$MSE_mean, r_c_771_ROR_p$MSE_mean),
  MSE_sd = c(r_b_771_p$MSE_sd, r_b_771_ROR_p$MSE_sd, r_c_771_p$MSE_sd, r_c_771_ROR_p$MSE_sd)
)
kable(data_table)
```

# Elastic Net - bootstrap
### 771 genes -> proliferation score (elastic Net - bootstrap)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/e_b_obj_771_p.RData")
e_b_771_p <- report02_out_ridge(e_b_obj_771_p)
```

### 771 genes -> ROR-proliferation score (elastic Net - bootstrap)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/e_b_obj_771_ROR_p.RData")
e_b_771_ROR_p <- report02_out_ridge(e_b_obj_771_ROR_p)
```

# Elastic Net: cross-validation
### 771 genes -> proliferation score  (elastic Net - repeated cross-validation)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/e_c_obj_771_prolif.RData")
e_c_771_p <- report02_out_ridge(e_c_obj_771_prolif)
```

### 771 genes -> ROR-proliferation score  (elastic Net - repeated cross-validation)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/e_c_obj_771_RORprolif.RData")
e_c_771_ROR_p <- report02_out_ridge(e_c_obj_771_RORprolif)
```

### Summery results: elastic net 771 genes bootstrap and repeated cross-validation
```{r}
data_table <- data.frame(
  Model = c("prolif boot", "ROR-prolif boot", "prolif rep cross-val", "ROR-prolif rep cross-val"),
  cor_mean = c(e_b_771_p$cor_mean, e_b_771_ROR_p$cor_mean, e_c_771_p$cor_mean, e_c_771_ROR_p$cor_mean),
  sd_cor = c(e_b_771_p$cor_sd, e_b_771_ROR_p$cor_sd, e_c_771_p$cor_sd, e_c_771_ROR_p$cor_sd),
  MSE_mean = c(e_b_771_p$MSE_mean, e_b_771_ROR_p$MSE_mean, e_c_771_p$MSE_mean, e_c_771_ROR_p$MSE_mean),
  MSE_sd = c(e_b_771_p$MSE_sd, e_b_771_ROR_p$MSE_sd, e_c_771_p$MSE_sd, e_c_771_ROR_p$MSE_sd)
)
kable(data_table)
```

# Boosting with stumps as base learner - bootstrap
### 771 genes -> proliferation score (boosting - bootstrap)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/xg_b_obj_771_p.RData")
xg_b_771_p <- report02_out_ridge(xg_b_obj_771_p)
```

### 771 genes -> ROR-proliferation score  (boosting - bootstrap)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/xg_b_obj_771_ROR_p.RData")
xg_b_771_ROR_p <- report02_out_ridge(xg_b_obj_771_ROR_p)
```

# Boosting with stumps as base learner cross-validation
### 771 genes -> proliferation score  (boosting - rep cross-val)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/xg_c_obj_771_prolif.RData")
xg_c_771_p <- report02_out_ridge(xg_c_obj_771_prolif)
```

### 771 genes -> ROR-proliferation score   (boosting - rep cross-val)
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/xg_c_obj_771_RORprolif.RData")
xg_c_771_ROR_p <- report02_out_ridge(xg_c_obj_771_RORprolif)
```

### Summery results: Boosting with stumps 771 genes bootstrap and repeated cross-validation
```{r}
data_table <- data.frame(
  Model = c("prolif boot", "ROR-prolif boot", "prolif rep cross-val", "ROR-prolif rep cross-val"),
  cor_mean = c(xg_b_771_p$cor_mean, xg_b_771_ROR_p$cor_mean, xg_c_771_p$cor_mean, xg_c_771_ROR_p$cor_mean),
  sd_cor = c(xg_b_771_p$cor_sd, xg_b_771_ROR_p$cor_sd, xg_c_771_p$cor_sd, xg_c_771_ROR_p$cor_sd),
  MSE_mean = c(xg_b_771_p$MSE_mean, xg_b_771_ROR_p$MSE_mean, xg_c_771_p$MSE_mean, xg_c_771_ROR_p$MSE_mean),
  MSE_sd = c(xg_b_771_p$MSE_sd, xg_b_771_ROR_p$MSE_sd, xg_c_771_p$MSE_sd, xg_c_771_ROR_p$MSE_sd)
)
kable(data_table)
```

# START USING DOMAIN KNOWLEDGE

# PCA on signature gene sets using repeated cross-validation

### Ridge: 771 genes -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/r_pca_c_771_RORprolif.RData")
r_pca_c_771_RORprolif <- report02_out_ridge(r_pca_c_771_RORprolif)
```

### Ridge: 771 genes -> ROR-proliferation score + interactions between PCs
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/r_pca_c_interact_771_RORprolif.RData")
r_pca_c_interact_771_RORprolif <- report02_out_ridge(r_pca_c_interact_771_RORprolif)
```

### Lasso: 771 genes -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/l_pca_c_771_RORprolif.RData")
l_pca_c_771_RORprolif <- report02_out(l_pca_c_771_RORprolif)
```

### Lasso: 771 genes -> ROR-proliferation score + interactions between PCs
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/l_pca_c_interact_771_RORprolif.RData")
l_pca_c_interact_771_RORprolif <- report02_out(l_pca_c_interact_771_RORprolif)
```

### ElasticNet: 771 genes -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/en_pca_c_771_RORprolif.RData")
en_pca_c_771_RORprolif <- report02_out(en_pca_c_771_RORprolif)
```

### ElasticNet: 771 genes -> ROR-proliferation score + interactions between PCs
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/en_pca_c_interact_771_RORprolif.RData")
en_pca_c_interact_771_RORprolif <- report02_out(en_pca_c_interact_771_RORprolif)
```

### Summery results: PCA ROR+proliferation score (repeated cross-validation) 
```{r}
data_table <- data.frame(
  Model = c("ridge", "ridge_interac", "lasso", "lasso_interact", "elastic", "elastic_interact"),
  cor_mean = c(r_pca_c_771_RORprolif$cor_mean, r_pca_c_interact_771_RORprolif$cor_mean, l_pca_c_771_RORprolif$cor_mean, l_pca_c_interact_771_RORprolif$cor_mean, en_pca_c_771_RORprolif$cor_mean, en_pca_c_interact_771_RORprolif$cor_mean),

  sd_cor = c(r_pca_c_771_RORprolif$cor_sd, r_pca_c_interact_771_RORprolif$cor_sd, l_pca_c_771_RORprolif$cor_sd, l_pca_c_interact_771_RORprolif$cor_sd, en_pca_c_771_RORprolif$cor_sd, en_pca_c_interact_771_RORprolif$cor_sd),

  MSE_mean = c(r_pca_c_771_RORprolif$MSE_mean, r_pca_c_interact_771_RORprolif$MSE_mean, l_pca_c_771_RORprolif$MSE_mean, l_pca_c_interact_771_RORprolif$MSE_mean, en_pca_c_771_RORprolif$MSE_mean, en_pca_c_interact_771_RORprolif$MSE_mean),

  MSE_sd = c(r_pca_c_771_RORprolif$MSE_sd, r_pca_c_interact_771_RORprolif$MSE_sd, l_pca_c_771_RORprolif$MSE_sd, l_pca_c_interact_771_RORprolif$MSE_sd, en_pca_c_771_RORprolif$MSE_sd, en_pca_c_interact_771_RORprolif$MSE_sd)
)
kable(data_table)
```

# Stacking on signature gene sets using repeated cross-validation

### Ridge: 771 genes -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/r_stack_c_771_RORprolif.RData")
r_stack_c_771_RORprolif <- report02_out_ridge(r_stack_c_771_RORprolif)
```

### Ridge: 771 genes -> ROR-proliferation score + interactions between PCs
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/r_stack_c_interact_771_RORprolif.RData")
r_stack_c_interact_771_RORprolif <- report02_out_ridge(r_stack_c_interact_771_RORprolif)
```

### Lasso: 771 genes -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/l_stack_c_771_RORprolif.RData")
l_stack_c_771_RORprolif <- report02_out(l_stack_c_771_RORprolif)
```

### Lasso: 771 genes -> ROR-proliferation score + interactions between PCs
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/l_stack_c_interact_771_RORprolif.RData")
l_stack_c_interact_771_RORprolif <- report02_out(l_stack_c_interact_771_RORprolif)
```

### ElasticNet: 771 genes -> ROR-proliferation score
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/en_stack_c_771_RORprolif.RData")
en_stack_c_771_RORprolif <- report02_out(en_stack_c_771_RORprolif)
```

### ElasticNet: 771 genes -> ROR-proliferation score + interactions between PCs
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/en_stack_c_interact_771_RORprolif.RData")
en_stack_c_interact_771_RORprolif <- report02_out(en_stack_c_interact_771_RORprolif)
```

### Summery results: Stacking ROR+proliferation score (repeated cross-validation) 
```{r}
data_table <- data.frame(
  Model = c("ridge", "ridge_interac", "lasso", "lasso_interact", "elastic", "elastic_interact"),
  cor_mean = c(r_stack_c_771_RORprolif$cor_mean, r_stack_c_interact_771_RORprolif$cor_mean, l_stack_c_771_RORprolif$cor_mean, l_stack_c_interact_771_RORprolif$cor_mean, en_stack_c_771_RORprolif$cor_mean, en_stack_c_interact_771_RORprolif$cor_mean),

  sd_cor = c(r_stack_c_771_RORprolif$cor_sd, r_stack_c_interact_771_RORprolif$cor_sd, l_stack_c_771_RORprolif$cor_sd, l_stack_c_interact_771_RORprolif$cor_sd, en_stack_c_771_RORprolif$cor_sd, en_stack_c_interact_771_RORprolif$cor_sd),

  MSE_mean = c(r_stack_c_771_RORprolif$MSE_mean, r_stack_c_interact_771_RORprolif$MSE_mean, l_stack_c_771_RORprolif$MSE_mean, l_stack_c_interact_771_RORprolif$MSE_mean, en_stack_c_771_RORprolif$MSE_mean, en_stack_c_interact_771_RORprolif$MSE_mean),

  MSE_sd = c(r_stack_c_771_RORprolif$MSE_sd, r_stack_c_interact_771_RORprolif$MSE_sd, l_stack_c_771_RORprolif$MSE_sd, l_stack_c_interact_771_RORprolif$MSE_sd, en_stack_c_771_RORprolif$MSE_sd, en_stack_c_interact_771_RORprolif$MSE_sd)
)
kable(data_table)
```

### Sparse Group Lasso: 771 genes -> ROR-proliferation score + interactions between PCs
```{r}
load("/Users/anders/Documents/MASTER/Cancer/R_codeP01/instances/sgl_c_771_RORprolif.RData")
sgl_c_771_RORprolif <- report02_out(sgl_c_771_RORprolif)
```

# Post Lasso
not done

# Summery of all results
### Summery results: lasso proliferation score (bootstrap) 
```{r}
library(knitr)
data_table <- tibble(
  Model = c("lasso 6 genes", "lasso 771 genes", "Nodes", "Residual additive", "Residual multiplicative"),
  cor_mean = c(l_b_6_p$cor_mean, l_b_771_p$cor_mean, l_b_nodes_p$cor_mean, l_b_residuals_771_p_a$cor_mean, l_b_residuals_771_p_m$cor_mean),
  sd_cor = c(l_b_6_p$cor_sd, l_b_771_p$cor_sd, l_b_nodes_p$cor_sd, l_b_residuals_771_p_a$cor_sd, l_b_residuals_771_p_m$cor_sd),
  MSE_mean = c(l_b_6_p$MSE_mean, l_b_771_p$MSE_mean, l_b_nodes_p$MSE_mean, l_b_residuals_771_p_a$MSE_mean, l_b_residuals_771_p_m$MSE_mean),
  MSE_sd = c(l_b_6_p$MSE_sd, l_b_771_p$MSE_sd, l_b_nodes_p$MSE_sd, l_b_residuals_771_p_a$MSE_sd, l_b_residuals_771_p_m$MSE_sd)
)
data_table |>  mutate(across(where(is.numeric), ~ round(., digits = 3))) |>  kable()
```

### Summery results: lasso ROR+proliferation score (bootstrap) 
```{r}
data_table <- tibble(
  Model = c("lasso 6 genes", "lasso 771 genes", "Nodes", "Residual additive", "Residual multiplicative"),
  cor_mean = c(l_b_6_ROR_p$cor_mean, l_b_771_ROR_p$cor_mean, l_b_nodes_ROR_p$cor_mean, l_b_residuals_ROR_p_a$cor_mean, l_b_residuals_ROR_p_m$cor_mean),
  sd_cor = c(l_b_6_ROR_p$cor_sd, l_b_771_ROR_p$cor_sd, l_b_nodes_ROR_p$cor_sd, l_b_residuals_ROR_p_a$cor_sd, l_b_residuals_ROR_p_m$cor_sd),
  MSE_mean = c(l_b_6_ROR_p$MSE_mean, l_b_771_ROR_p$MSE_mean, l_b_nodes_ROR_p$MSE_mean, l_b_residuals_ROR_p_a$MSE_mean, l_b_residuals_ROR_p_m$MSE_mean),
  MSE_sd = c(l_b_6_ROR_p$MSE_sd, l_b_771_ROR_p$MSE_sd, l_b_nodes_ROR_p$MSE_sd, l_b_residuals_ROR_p_a$MSE_sd, l_b_residuals_ROR_p_m$MSE_sd)
)
data_table |>  mutate(across(where(is.numeric), ~ round(., digits = 3))) |>  kable()
```

### Summery results: lasso proliferation score (repeated cross-validation) 
```{r}
library(knitr)
data_table <- tibble(
  Model = c("lasso 6 genes", "lasso 771 genes", "Nodes", "Residual additive", "Residual multiplicative"),
  cor_mean = c(l_c_6_p$cor_mean, l_c_771_p$cor_mean, l_c_nodes_p$cor_mean, l_c_residuals_771_p_a$cor_mean, l_c_residuals_771_p_m$cor_mean),
  sd_cor = c(l_c_6_p$cor_sd, l_c_771_p$cor_sd, l_c_nodes_p$cor_sd, l_c_residuals_771_p_a$cor_sd, l_c_residuals_771_p_m$cor_sd),
  MSE_mean = c(l_c_6_p$MSE_mean, l_b_771_p$MSE_mean, l_c_nodes_p$MSE_mean, l_c_residuals_771_p_a$MSE_mean, l_c_residuals_771_p_m$MSE_mean),
  MSE_sd = c(l_c_6_p$MSE_sd, l_c_771_p$MSE_sd, l_c_nodes_p$MSE_sd, l_c_residuals_771_p_a$MSE_sd, l_c_residuals_771_p_m$MSE_sd)
)
data_table |>  mutate(across(where(is.numeric), ~ round(., digits = 3))) |>  kable()
```

### Summery results: lasso ROR+proliferation score (repeated cross-validation) 
```{r}
data_table <- tibble(
  Model = c("lasso 6 genes", "lasso 771 genes", "Nodes", "Residual additive", "Residual multiplicative"),
  cor_mean = c(l_c_6_ROR_p$cor_mean, l_c_771_ROR_p$cor_mean, l_c_nodes_ROR_p$cor_mean, l_c_residuals_ROR_p_a$cor_mean, l_c_residuals_ROR_p_m$cor_mean),
  sd_cor = c(l_c_6_ROR_p$cor_sd, l_c_771_ROR_p$cor_sd, l_c_nodes_ROR_p$cor_sd, l_c_residuals_ROR_p_a$cor_sd, l_c_residuals_ROR_p_m$cor_sd),
  MSE_mean = c(l_c_6_ROR_p$MSE_mean, l_c_771_ROR_p$MSE_mean, l_c_nodes_ROR_p$MSE_mean, l_c_residuals_ROR_p_a$MSE_mean, l_c_residuals_ROR_p_m$MSE_mean),
  MSE_sd = c(l_c_6_ROR_p$MSE_sd, l_c_771_ROR_p$MSE_sd, l_c_nodes_ROR_p$MSE_sd, l_c_residuals_ROR_p_a$MSE_sd, l_c_residuals_ROR_p_m$MSE_sd)
)
data_table |>  mutate(across(where(is.numeric), ~ round(., digits = 3))) |>  kable()
```

### Summery results: ridge 771 genes bootstrap and repeated cross-validation
```{r}
data_table <- tibble(
  Model = c("prolif boot", "ROR-prolif boot", "prolif rep cross-val", "ROR-prolif rep cross-val"),
  cor_mean = c(r_b_771_p$cor_mean, r_b_771_ROR_p$cor_mean, r_c_771_p$cor_mean, r_c_771_ROR_p$cor_mean),
  sd_cor = c(r_b_771_p$cor_sd, r_b_771_ROR_p$cor_sd, r_c_771_p$cor_sd, r_c_771_ROR_p$cor_sd),
  MSE_mean = c(r_b_771_p$MSE_mean, r_b_771_ROR_p$MSE_mean, r_c_771_p$MSE_mean, r_c_771_ROR_p$MSE_mean),
  MSE_sd = c(r_b_771_p$MSE_sd, r_b_771_ROR_p$MSE_sd, r_c_771_p$MSE_sd, r_c_771_ROR_p$MSE_sd)
)
data_table |>  mutate(across(where(is.numeric), ~ round(., digits = 3))) |>  kable()
```

### Summery results: elastic net 771 genes bootstrap and repeated cross-validation
```{r}
data_table <- tibble(
  Model = c("prolif boot", "ROR-prolif boot", "prolif rep cross-val", "ROR-prolif rep cross-val"),
  cor_mean = c(e_b_771_p$cor_mean, e_b_771_ROR_p$cor_mean, e_c_771_p$cor_mean, e_c_771_ROR_p$cor_mean),
  sd_cor = c(e_b_771_p$cor_sd, e_b_771_ROR_p$cor_sd, e_c_771_p$cor_sd, e_c_771_ROR_p$cor_sd),
  MSE_mean = c(e_b_771_p$MSE_mean, e_b_771_ROR_p$MSE_mean, e_c_771_p$MSE_mean, e_c_771_ROR_p$MSE_mean),
  MSE_sd = c(e_b_771_p$MSE_sd, e_b_771_ROR_p$MSE_sd, e_c_771_p$MSE_sd, e_c_771_ROR_p$MSE_sd)
)
data_table |>  mutate(across(where(is.numeric), ~ round(., digits = 3))) |>  kable()
```

### Summery results: Boosting with stumps 771 genes bootstrap and repeated cross-validation
```{r}
data_table <- tibble(
  Model = c("prolif boot", "ROR-prolif boot", "prolif rep cross-val", "ROR-prolif rep cross-val"),
  cor_mean = c(xg_b_771_p$cor_mean, xg_b_771_ROR_p$cor_mean, xg_c_771_p$cor_mean, xg_c_771_ROR_p$cor_mean),
  sd_cor = c(xg_b_771_p$cor_sd, xg_b_771_ROR_p$cor_sd, xg_c_771_p$cor_sd, xg_c_771_ROR_p$cor_sd),
  MSE_mean = c(xg_b_771_p$MSE_mean, xg_b_771_ROR_p$MSE_mean, xg_c_771_p$MSE_mean, xg_c_771_ROR_p$MSE_mean),
  MSE_sd = c(xg_b_771_p$MSE_sd, xg_b_771_ROR_p$MSE_sd, xg_c_771_p$MSE_sd, xg_c_771_ROR_p$MSE_sd)
)
data_table |>  mutate(across(where(is.numeric), ~ round(., digits = 3))) |>  kable()
```

## Summary using domain knowledge

### Summery results: PCA ROR+proliferation score (repeated cross-validation) 
```{r}
data_table <- tibble(
  Model = c("ridge", "ridge_interac", "lasso", "lasso_interact", "elastic", "elastic_interact"),
  cor_mean = c(r_pca_c_771_RORprolif$cor_mean, r_pca_c_interact_771_RORprolif$cor_mean, l_pca_c_771_RORprolif$cor_mean, l_pca_c_interact_771_RORprolif$cor_mean, en_pca_c_771_RORprolif$cor_mean, en_pca_c_interact_771_RORprolif$cor_mean),

  sd_cor = c(r_pca_c_771_RORprolif$cor_sd, r_pca_c_interact_771_RORprolif$cor_sd, l_pca_c_771_RORprolif$cor_sd, l_pca_c_interact_771_RORprolif$cor_sd, en_pca_c_771_RORprolif$cor_sd, en_pca_c_interact_771_RORprolif$cor_sd),

  MSE_mean = c(r_pca_c_771_RORprolif$MSE_mean, r_pca_c_interact_771_RORprolif$MSE_mean, l_pca_c_771_RORprolif$MSE_mean, l_pca_c_interact_771_RORprolif$MSE_mean, en_pca_c_771_RORprolif$MSE_mean, en_pca_c_interact_771_RORprolif$MSE_mean),

  MSE_sd = c(r_pca_c_771_RORprolif$MSE_sd, r_pca_c_interact_771_RORprolif$MSE_sd, l_pca_c_771_RORprolif$MSE_sd, l_pca_c_interact_771_RORprolif$MSE_sd, en_pca_c_771_RORprolif$MSE_sd, en_pca_c_interact_771_RORprolif$MSE_sd)
)
data_table |>  mutate(across(where(is.numeric), ~ round(., digits = 3))) |>  kable()
```

### Summery results: Stacking ROR+proliferation score (repeated cross-validation) 
```{r}
data_table <- tibble(
  Model = c("ridge", "ridge_interac", "lasso", "lasso_interact", "elastic", "elastic_interact"),
  cor_mean = c(r_stack_c_771_RORprolif$cor_mean, r_stack_c_interact_771_RORprolif$cor_mean, l_stack_c_771_RORprolif$cor_mean, l_stack_c_interact_771_RORprolif$cor_mean, en_stack_c_771_RORprolif$cor_mean, en_stack_c_interact_771_RORprolif$cor_mean),

  sd_cor = c(r_stack_c_771_RORprolif$cor_sd, r_stack_c_interact_771_RORprolif$cor_sd, l_stack_c_771_RORprolif$cor_sd, l_stack_c_interact_771_RORprolif$cor_sd, en_stack_c_771_RORprolif$cor_sd, en_stack_c_interact_771_RORprolif$cor_sd),

  MSE_mean = c(r_stack_c_771_RORprolif$MSE_mean, r_stack_c_interact_771_RORprolif$MSE_mean, l_stack_c_771_RORprolif$MSE_mean, l_stack_c_interact_771_RORprolif$MSE_mean, en_stack_c_771_RORprolif$MSE_mean, en_stack_c_interact_771_RORprolif$MSE_mean),

  MSE_sd = c(r_stack_c_771_RORprolif$MSE_sd, r_stack_c_interact_771_RORprolif$MSE_sd, l_stack_c_771_RORprolif$MSE_sd, l_stack_c_interact_771_RORprolif$MSE_sd, en_stack_c_771_RORprolif$MSE_sd, en_stack_c_interact_771_RORprolif$MSE_sd)
)
data_table |>  mutate(across(where(is.numeric), ~ round(., digits = 3))) |>  kable()
```

# Summery most interesting maybe
Response is just ROR+proliferation score, and only used repeated cross-validation.

```{r}
# Create a data frame
data_table <- tibble(
  Model = c("Lasso", "Ridge", "ElasticNet", "Boosting", "Residual (lasso/additive)", "Sparse group lasso",
            "PCA ON GENE SETS", "ridge", "ridge_interac", "lasso", "lasso_interact", "elastic", "elastic_interact", 
            "STACKING ON GENE SETS", "ridge", "ridge_interac", "lasso", "lasso_interact", "elastic", "elastic_interact"),

  cor_mean = c(
    l_c_771_ROR_p$cor_mean,
    r_c_771_ROR_p$cor_mean,
    e_c_771_ROR_p$cor_mean,
    xg_c_771_ROR_p$cor_mean,
    l_c_residuals_ROR_p_a$cor_mean,
    sgl_c_771_RORprolif$cor_mean,
    NA,
    r_pca_c_771_RORprolif$cor_mean, 
    r_pca_c_interact_771_RORprolif$cor_mean, 
    l_pca_c_771_RORprolif$cor_mean, 
    l_pca_c_interact_771_RORprolif$cor_mean, 
    en_pca_c_771_RORprolif$cor_mean, 
    en_pca_c_interact_771_RORprolif$cor_mean,
    NA, 
    r_stack_c_771_RORprolif$cor_mean, 
    r_stack_c_interact_771_RORprolif$cor_mean, 
    l_stack_c_771_RORprolif$cor_mean, 
    l_stack_c_interact_771_RORprolif$cor_mean, 
    en_stack_c_771_RORprolif$cor_mean,
    en_stack_c_interact_771_RORprolif$cor_mean),

  sd_cor = c(
    l_c_771_ROR_p$cor_sd,
    r_c_771_ROR_p$cor_sd,
    e_c_771_ROR_p$cor_sd,
    xg_c_771_ROR_p$cor_sd,
    l_c_residuals_ROR_p_a$cor_sd,
    sgl_c_771_RORprolif$cor_sd,
    NA,
    r_pca_c_771_RORprolif$cor_sd, 
    r_pca_c_interact_771_RORprolif$cor_sd, 
    l_pca_c_771_RORprolif$cor_sd, 
    l_pca_c_interact_771_RORprolif$cor_sd, 
    en_pca_c_771_RORprolif$cor_sd, 
    en_pca_c_interact_771_RORprolif$cor_sd,
    NA,
    r_stack_c_771_RORprolif$cor_sd, 
    r_stack_c_interact_771_RORprolif$cor_sd, 
    l_stack_c_771_RORprolif$cor_sd, 
    l_stack_c_interact_771_RORprolif$cor_sd, 
    en_stack_c_771_RORprolif$cor_sd, 
    en_stack_c_interact_771_RORprolif$cor_sd),

  MSE_mean = c(    
    l_c_771_ROR_p$MSE_mean,
    r_c_771_ROR_p$MSE_mean,
    e_c_771_ROR_p$MSE_mean,
    xg_c_771_ROR_p$MSE_mean,
    l_c_residuals_ROR_p_a$MSE_mean,
    sgl_c_771_RORprolif$MSE_mean,
    NA, 
    r_pca_c_771_RORprolif$MSE_mean, 
    r_pca_c_interact_771_RORprolif$MSE_mean, 
    l_pca_c_771_RORprolif$MSE_mean, 
    l_pca_c_interact_771_RORprolif$MSE_mean, 
    en_pca_c_771_RORprolif$MSE_mean, 
    en_pca_c_interact_771_RORprolif$MSE_mean,
    NA,
    r_stack_c_771_RORprolif$MSE_mean, 
    r_stack_c_interact_771_RORprolif$MSE_mean, 
    l_stack_c_771_RORprolif$MSE_mean,
    l_stack_c_interact_771_RORprolif$MSE_mean, 
    en_stack_c_771_RORprolif$MSE_mean, 
    en_stack_c_interact_771_RORprolif$MSE_mean),

  MSE_sd = c(    
    l_c_771_ROR_p$MSE_sd,
    r_c_771_ROR_p$MSE_sd,
    e_c_771_ROR_p$MSE_sd,
    xg_c_771_ROR_p$MSE_sd,
    l_c_residuals_ROR_p_a$MSE_sd,
    sgl_c_771_RORprolif$MSE_sd,
    NA,
    r_pca_c_771_RORprolif$MSE_sd, 
    r_pca_c_interact_771_RORprolif$MSE_sd, 
    l_pca_c_771_RORprolif$MSE_sd, 
    l_pca_c_interact_771_RORprolif$MSE_sd, 
    en_pca_c_771_RORprolif$MSE_sd, 
    en_pca_c_interact_771_RORprolif$MSE_sd,
    NA,
    r_stack_c_771_RORprolif$MSE_sd,
    r_stack_c_interact_771_RORprolif$MSE_sd, 
    l_stack_c_771_RORprolif$MSE_sd, 
    l_stack_c_interact_771_RORprolif$MSE_sd, 
    en_stack_c_771_RORprolif$MSE_sd, 
    en_stack_c_interact_771_RORprolif$MSE_sd)
)

# Create a custom function to replace NAs with empty strings
replace_na_with_blank <- function(x) {
  if (is.numeric(x)) {
    ifelse(is.na(x), "", round(x, digits = 3))
  } else {
    x
  }
}

# Apply the custom function to the data_table
data_table <- data_table %>%
  mutate(across(everything(), replace_na_with_blank))

data_table %>% kable()

#data_table <- data_table |> mutate(across(where(is.numeric), ~ round(., digits = 3))) 
#data_table |>  kable()

```